Reproducibility in plant pathology: where do we stand and a way forward

### 2022-04-24 analysis/paper/paper.Rmd
correct number of articles evaluated

Open research practices have been highlighted extensively during the last ten years in many fields of scientific study as essential standards needed to promote transparency and reproducibility of scientific results.
Scientific claims can only be evaluated based on how protocols, materials, equipment and methods were described; data were collected and prepared; and, analyses were conducted.
Openly sharing protocols, data and computational code is central for current scholarly dissemination and communication, but in many fields, including plant pathology, adoption of these practices has been slow.
We randomly selected 300 articles published from 2012 to 2018 across 21 journals representative of the plant pathology discipline and assigned them scores reflecting their openness and reproducibility.
We found that most of the articles were not following protocols for open science, and were failing to share data or code in a reproducible way.
We also propose that use of open-source tools facilitates reproducible work and analyses benefitting not just readers, but the authors as well.
Finally, we also provide ideas and tools to promote open, reproducible research practices among plant pathologists.

@@

It goes without saying that plant pathologists already provide information on protocols and chemicals allowing for reproducibility.
However, biological specimens such as strains, cultures or cultivars are not available.
These cases do constitute a lack of reproducibility but will not be covered here.

@@

## Article Selection and Evaluation Events

In an effort to understand where as a discipline we stand as plant pathologists regarding open science and reproducible research, we surveyed a broad selection of articles to represent a broad swathe of publications to evaluate our status.
We hand-picked 21 journals which we felt represented research publications in the field of plant pathology (Table \@ref(tab:journals)) that encompassed a range of subject matter foci, applied and fundamental work, country of origin and ranking metrics, _e.g._ quartile range or citation index, to represent the discipline.
The aim was to gather as complete an overview of the status of reproducibility in plant pathology journals as possible and avoid bias in the findings by skewing towards high-impact journals that may have a greater influence.
From those 21 journals we randomly selected 450 articles published from 2012 to 2021.
Using R [@RCT2022], two lists were created, the first was a list of the 21 journals and the second was a list of the evaluators that were evaluating the articles for reproducibility.
Initially there were four evaluators, later a fifth was added and the list was recreated.
There were three evaluation events that took place to increase the yearly coverage of the evaluations.
Each evaluator was assigned 50 articles to evaluate during an evaluation event.
A list of randomly generated numbers to represent page numbers from one to 150, sampled with replacement, was assigned to a randomized list of the 21 journals for .
This was done because some journals restart their numbering with each issue and also ensures that the journal is more likely to have a page number corresponding to the randomly generated value.
This also assumes that there is no effect or bias on reproducibility based on the time of year that an article was published since most journals start with page number one at the beginning of the year.

Articles were then manually selected by visiting the journals' websites and selecting the articles within which the randomly assigned page numbers fell, _i.e._, if the page number was 32, the article that started on page 28 and ended on page 35 was selected as it contained page 32.
In cases where an article was not suitable, _e.g._, a review or otherwise not related to plant pathology or the randomly assigned numbers for that journal and year fell within the same article, the next article in that journal was selected until a suitable article was found.

The each of the five authors were each assigned to rate a randomised list of journal articles using scoring criteria devised for the purposes of this research (Table \@ref(tab:scoring)) in three different scoring events as their time allowed for their participation.
In the first event, three authors (AS, EDP, and ZF) evaluated 200 articles, in the second event two authors evaluated 100 articles (EDP, KA) and in the third round, five authors (AS, EDP, KA, ZF, NG) evaluated 150 articles for a total of 450 articles.

@@

Each journal was classified as to whether they were completely open (TRUE), behind a paywall (FALSE) or a combination (BOTH).
The five-year impact factor for 2022 for each journal was retrieved from InCites Journal Citation Reports, Clarivate Analytics and entered in a separate sheet in the Google sheets file.
This was downloaded, saved as an Open Document Spreadsheet (ODS) file and is left-joined with the article notes using a `left_join()` from the 'dplyr' package [@Wickham2020] for analysis when a custom function for this analysis, `import_notes()`, was executed.

@@

Values for software packages were checked for spelling consistency and corrections were made manually where necessary.
When working with these data in R, all strings of software character values were converted to fully upper-case to standardise the capitalisation and alleviate any issues with capitalisation used between evaluators.
A custom function, `import_notes()`, was written to import the data, format the columns properly and calculate the overall reproducibility score in R [@RCT2022].

@@

The statistical analysis was performed using R version `r version$version.string` [@RCT2022] on an Apple MacBook Air (M1, 2020).
Bayesian mixed-effect models were fit using the contributed package 'brms' (version `r packageVersion("brms")`) [@Burkner2017; @Burkner2018; @Burkner2021] to evaluate the fixed effects of the publication (journal title), article class, molecular focus, five-year journal impact factor, journal open access policy on the reproducibility index and year to test for changes over time in the reproducibility index.

@@

Model finesses were evaluated using model summaries and diagnostic plots from 'brms' 'brms' (version `r packageVersion("brms")`) [@Burkner2017; @Burkner2018; @Burkner2021] and posterior fits using `pp_check()` (Figure \@ref(fig:posteriors)) from the contributed R package 'bayesplot' (version `r packageVersion("bayesplot")`) [@Gabry2019].

@@

Data were visualised using 'ggplot2' ((version `r packageVersion("ggplot2")`)) [@Wickham2016], 'ggpubr' ((version `r packageVersion("ggpubr")`)) [@Kassambara2020] and patchwork ((version `r packageVersion("patchwork")`)) [@Pedersen2020].

All methods are described in greater detail with the code necessary to reproduce the work in @Sparks2022.

@@

The fourth level includes packaging the analysis as a software package, _e.g._, R, Julia or Python package at a minimum, with the use of continuous integration and possibly a software container like Docker as @Khaliq2020 have done.
Their research compendium provides a Docker instance, DOI and full R package that enable readers to fully replicate their analysis of _Ascochyta rabiei_ conidia dispersal in chickpea using the data collected and stepping through other points where weather data were investigated and various models were fit before deciding on the best fit and recreate any figures as published in the article [@Khaliq2020a].
When this level is employed, tools such as Binder, [https://mybinder.org](https://mybinder.org), can be used that allow readers and reviewers to launch an interactive session in their web browser and interact with the data and rerun the analysis in an RStudio instance or Jupyter notebooks as Mioroni et al. used [-@Miorini2018, [-@Miorini2019]].
Kamvar et al. [-@Kamvar2015] took a slightly different approach by including all files necessary for the analysis and most output files in a repository [@Kamvar2014] that also included an installable R package that was used for the original analysis.
While there are many other methods, these two approaches illustrate some of the best-practices where the data and other files were deposited in repositories with DOIs and reproducibility issues were addressed by using R packages to handle dependencies and other versioning issues making the work more portable.

@@

The use of programming or scripting languages such as R, Julia, SAS or Python enables you as a scientist to keep very detailed records of what was computationally performed.
This is as opposed to using software such as spreadsheet programs like Excel, Google Sheets, Numbers, Calc or others which can be used for simple statistical analyses and visualization or other point-and-click software packages that do not enable you to keep an accurate record of the steps taken to import, format, visualise and analyse data.

@@

A.H.S, E.DP., Z.F. and N.J.G. conceived of the presented idea and scoring system.
E.DP. provided five-year impact factors.
A.H.S., E.DP., Z.F. K.A. and N.G. evaluated articles for scoring.
A.H.S. designed the computational framework for analysis and created the research compendium with feedback from E.DP. and Z.F.
A.H.S. wrote the manuscript in consultation with E.DP.
All authors provided critical feedback and helped contribute to the final version of the manuscript.

@@

```{r posteriors, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Posterior distribution visualisations for each of six models fitted to scoring data that were used to evaluate factors on reproducibility of papers published in 20 plant pathology journals or plant pathology focused articles from other specialised journals. Six factors were evaluated, (A) Publication (Journal Title), (B) Article Class, (C) Molecular Focus, (D) Five Year Impact Factor of Journal, (E) Journal Open Access Policy and (F) Year of Publication where y is the data's density and y~rep~ are ten random posterior draws. Model fitnesses were found to be good for all models and no effects of any of the factors were observed on the reproducibility score.", cache=FALSE, fig.env='figure', out.width=0, out.height=0}

@@

p <- p + plot_annotation(tag_levels = "A",
                         title = "Bayesian posterior distributions of evaluation score models"
                         )

ggsave(filename = "../figures/Sparks et al Figure 1.eps",
       plot = p,
       width = 6,
       height = 6,
       units = "in",
       device = cairo_ps)
extrafont::embed_fonts(
   file = ("../figures/Sparks et al Figure 1.eps"),
   outfile = ("../figures/Sparks et al Figure 1.eps"),
   options = "-dEPSCrop"
)
knitr::include_graphics(path = "../figures/Sparks et al Figure 1.eps")
```

@@

# put them all together
p <- (A | B) / (C | D)
p <- p + plot_annotation(tag_levels = "A",
                         title = "Aggregated article scores")
ggsave(filename = "../figures/Sparks et al Figure 2.eps",
       plot = p,
       width = 6,
       height = 6,
       units = "in",
       device = cairo_ps)
extrafont::embed_fonts(
   file = ("../figures/Sparks et al Figure 2.eps"),
   outfile = ("../figures/Sparks et al Figure 2.eps"),
   options = "-dEPSCrop"
)
knitr::include_graphics(path = "../figures/Sparks et al Figure 2.eps")
```

@@

    node[shape = plaintext, width = 1.2, fontname = Helvetica, fontsize = 20]
    'An open and reproducible workflow'

    node [shape = oval, style = filled, fillcolor = grey99, width = 1.2, fontname = Helvetica]
    Question

    node [shape = box, style = filled, fillcolor = grey99, width = 1.2, fontname = Helvetica]
    Methodology; Data; Analysis; Manuscript;

    node [shape = box, style = filled, fillcolor = grey90, width = 1.2, fontname = Helvetica]
    Description; Availability; Citation

@@

    node [shape = box, style = filled, fillcolor = grey70, width = 1.2]
    'Source Code'; 'Data File'; Metadata; 'Public Repository';

    Question->Methodology->Data->Analysis
    Analysis->Manuscript
    Methodology->Description
    Methodology->Citation
    Methodology->Availability
    Manuscript->Supplement
    Data->'Data File'->'Public Repository'
    Data->Metadata->'Public Repository'
    Analysis->'Binary Code'->Supplement
    Analysis->'Source Code'->'Public Repository'
    Data->'Binary File'->Supplement
  }",
  width = 1000,
  height = 750
))
rsvg_ps(svg = charToRaw(p), file = "../figures/Sparks et al Figure 3.eps")
extrafont::embed_fonts(
   file = ("../figures/Sparks et al Figure 3.eps"),
   outfile = ("../figures/Sparks et al Figure 3.eps"),
   options = "-dEPSCrop"
)
knitr::include_graphics(path = "../figures/Sparks et al Figure 3.eps")
```

>>>>

Open research practices have been highlighted extensively during the last ten years in many fields of scientific study as essential standards needed to promote transparency and reproducibility of scientific results.
Scientific claims can only be evaluated based on how protocols, materials, equipment and methods were described; data were collected and prepared; and, analyses were conducted.
Openly sharing protocols, data and computational code is central for current scholarly dissemination and communication, but in many fields, including plant pathology, adoption of these practices has been slow.
We randomly selected 450 articles published from 2012 to 2018 across 21 journals representative of the plant pathology discipline and assigned them scores reflecting their openness and reproducibility.
We found that most of the articles were not following protocols for open science, and were failing to share data or code in a reproducible way.
We also propose that use of open-source tools facilitates reproducible work and analyses benefitting not just readers, but the authors as well.
Finally, we also provide ideas and tools to promote open, reproducible research practices among plant pathologists.

@@

It goes without saying that plant pathologists already provide information on protocols and chemicals allowing for reproducibility.
However, very often biological specimens such as strains, cultures or cultivars are often not available.
These cases do constitute a lack of reproducibility but will not be covered here.

@@

## Article Selection

In an effort to understand where as a discipline we stand as plant pathologists in regards to open science and reproducible research, we surveyed a broad selection of articles to represent a broad swathe of publications to evaluate our status.
We randomly selected 450 articles published from 2012 to 2018 across 21 journals (Table \@ref(tab:journals)) and assigned them scores reflecting their openness and reproducibility.
The journals represented both fundamental and applied areas as well as pathogen or other group-specific publications.
A list of randomly selected pages was assigned to a randomized list of the 21 journals where the page number fell within an article for the given journal.
In cases where an article was not suitable, _e.g._, a review or otherwise not related to plant pathology, the next article was selected until a suitable article was found.
Notes regarding the selection of articles can be found in the file, "Reproducibility_in_plant_pathology_notes.ods", available in this paper's repository in `/inst/extdata'.
The pages list was numbered from page one and went to page 150.
This was done since some journals restart their numbering with each issue and also ensures that the journal is more likely to have a page number corresponding to the randomly generated value.
This also assumes that there is no effect or bias on reproducibility based on the time of year that an article was published since most journals start with page number one at the beginning of the year.

@@

Four individuals were assigned to rate a randomised list of journal articles using scoring criteria devised for the purposes of this research (Table \@ref(tab:scoring)).

Each journal was classified as to whether they were completely open (TRUE), behind a paywall (FALSE) or a combination (BOTH).
The five-year impact factor for 2018 for each journal was retrieved from InCites Journal Citation Reports, Clarivate Analytics and entered in a separate sheet in the Google sheets file.
This was downloaded, saved as an Open Document Spreadsheet (ODS) file and is left-joined with the article notes using a `left_join()` from the 'dplyr' package [@Wickham2020] for analysis when the `import_notes()` command is executed.

@@

Values for software packages were checked for spelling consistency and corrections were made manually where necessary.
When working with these data in R, all strings of software character values were converted to fully upper-case to standardise the capitalisation and alleviate any issues with capitalisation used between evaluators.
A custom R function, `import_notes()`, was written to import the data, format the columns properly and calculate the overall reproducibility score in R [@RCT2021].

@@

The statistical analysis was performed using R version 4.1.0 [@RCT2021] on an Apple MacBook Air (M1, 2020).
Bayesian mixed-effect models were fit using the contributed package 'brms' (version 2.15.0) [@Burkner2018] to evaluate the fixed effects of the publication (journal title), article class, molecular focus, five-year journal impact factor, journal open access policy on the reproducibility index and year to test for changes over time in the reproducibility index.

@@

Model finesses were evaluated using model summaries and diagnostic plots from 'brms' (version 2.15.0) [@Burkner2018] and posterior fits using `pp_check()`  (Figure \@ref(fig:posteriors)) from the contributed R package 'bayesplot' (version 1.8.1) [@Gabry2019].

@@

Data were visualised using 'ggplot2' (version 3.3.5) [@Wickham2016], 'ggpubr' (version 0.4.0) [@Kassambara2020] and patchwork (version 1.1.1) [@Pedersen2020].

@@

The fourth level includes packaging the analysis as a software package, _e.g._, R, Julia or Python package at a minimum, with the use of continuous integration and possibly a software container like Docker as Khaliq et al. [@Khaliq2020] have done.
Their research compendium provides a Docker instance, DOI and full R package that enable readers to fully replicate their analysis of _Ascochyta rabiei_ conidia dispersal in chickpea using the data collected and stepping through other points where weather data were investigated and various models were fit before deciding on the best fit and recreate any figures as published in the article [@Khaliq2020a].
When this level is employed, tools such as Binder, [https://mybinder.org](https://mybinder.org), can be used that allow readers and reviewers to launch an interactive session in their web browser and interact with the data and rerun the analysis in an RStudio instance or Jupyter notebooks as Mioroni et al. used [-@Miorini2018, [-@Miorini2019]].
Kamvar et al. [-@Kamvar2015] took a slightly different approach by including all files necessary for the analysis and most output files in a repository [@Kamvar2014] that also included an installable R package that was used for the original analysis.
While there are many other methods, these two approaches illustrate some of the best-practices where the data and other files were deposited in repositories with DOIs and reproducibility issues were addressed by using R packages to handle dependencies and other versioning issues making the work more portable.

@@

The use of programming or scripting languages such as R, Julia, SAS or Python enables you as a scientist to keep very detailed records of what was computationally performed.
This is as opposed to using software such as spreadsheet programs like Excel, Google Sheets, Numbers, Calc or others which can be used for simple statistical analyses and visualization (but should not be) or other point-and-click software packages that do not enable you to keep an accurate record of the steps taken to import, format, visualise and analyse data.

@@

A.H.S, E.DP., Z.F. and N.J.G. conceived of the presented idea and scoring system.
N.J.G. provided five-year impact factors.
A.H.S., E.DP., Z.F. and K.dSA. evaluated articles for scoring.
A.H.S. designed the computational framework for analysis and created the research compendium with feedback from E.DP. and Z.F.
A.H.S. wrote the manuscript in consultation with E.DP.
All authors provided critical feedback and helped contribute to the final version of the manuscript.

@@

```{r posteriors, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Posterior distribution visualisations for each of six models fitted to scoring data that were used to evaluate the effects of factors affecting the reproducibility of papers published in 20 plant pathology journals or plant pathology focused articles from other specialised journals. Six factors were evaluated, (A) Publication (Journal Title), (B) Article Class, (C) Molecular Focus, (D) Five Year Impact Factor of Journal, (E) Journal Open Access Policy and (F) Year of Publication where y is the data's density and y~rep~ are ten random posterior draws. Model fitnesses were found to be good for all models and no effects of any of the factors were observed on the reproducibility score.", cache=FALSE, fig.env='figure', out.width=0, out.height=0}

@@

p <- p + plot_annotation(tag_levels = "A")
ggsave(filename = "../figures/Sparks et al Figure 1.eps",
       plot = p,
       width = 6,
       height = 6,
       units = "in",
       device = cairo_ps)
extrafont::embed_fonts(
   file = ("../figures/Sparks et al Figure 1.eps"),
   outfile = ("../figures/Sparks et al Figure 1.eps"),
   options = "-dEPSCrop"
)
knitr::include_graphics(path = "../figures/Sparks et al Figure 1.eps")
```

@@

# put them all together
p <- (A | B) / (C | D)
p <- p + plot_annotation(tag_levels = "A")
ggsave(filename = "../figures/Sparks et al Figure 2.eps",
       plot = p,
       width = 6,
       height = 6,
       units = "in",
       device = cairo_ps)
extrafont::embed_fonts(
   file = ("../figures/Sparks et al Figure 2.eps"),
   outfile = ("../figures/Sparks et al Figure 2.eps"),
   options = "-dEPSCrop"
)
knitr::include_graphics(path = "../figures/Sparks et al Figure 2.eps")
```

@@

    node [shape = box, style = filled, fillcolor = grey99, width = 1.2,fontname = Helvetica]
    Question; Methodology; Data; Analysis; Manuscript;

    node [shape = box, style = filled, fillcolor = grey90, width = 1.2,fontname = Helvetica]
    Description; Availability; Citation

@@

    node [shape = box, style = filled, fillcolor = grey70, width = 1.2]
    'Source Code'; 'Data File'; 'Public Repository';

    Question->Methodology->Data->Analysis
    Analysis->Manuscript
    Methodology->Description
    Methodology->Citation
    Methodology->Availability
    Manuscript->Supplement
    Data->'Data File'->'Public Repository'
    Analysis->'Binary Code'->Supplement
    Analysis->'Source Code'->'Public Repository'
    Data->'Binary File'->Supplement
  }",
  width = 1000,
  height = 750
))
rsvg_ps(svg = charToRaw(p), file = "../figures/Sparks et al Figure 3.eps")
extrafont::embed_fonts(
   file = ("../figures/Sparks et al Figure 3.eps"),
   outfile = ("../figures/Sparks et al Figure 3.eps"),
   options = "-dEPSCrop"
)
knitr::include_graphics(path = "../figures/Sparks et al Figure 3.eps")
```
